---
title: "p8105_hw5_co2554"
author: "Camille Okonkwo"
output: github_document
---

```{r setup, include = FALSE, message = FALSE, warning = FALSE}
library(tidyverse)
library(broom)

knitr::opts_chunk$set(
  echo = TRUE, 
  warning = FALSE,
  fig.width = 6, 
  fig.asp = .6, 
  out.width = "90%"
  )

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

Set seed for reproducibility.

```{r}
set.seed(12345)
```

### Problem 2

This zip file contains data from a longitudinal study that included a control arm and an experimental arm. Data for each participant is included in a separate file, and file names include the subject ID and arm.

Create a tidy dataframe containing data from all participants, including the subject ID, arm, and observations over time:

```{r import_tidy}
long_study_files = list.files(path = "data/", pattern = ".csv", full.names = TRUE)

long_study_df =
  long_study_files |> 
  map_df(~{
    read.csv(.x)  |> 
      janitor::clean_names()  |> 
      mutate(
        subject_id = as.numeric(str_extract(.x, "\\d+")),
        arm = ifelse(grepl("con", .x), "Control", "Experimental")
      )  |> 
      pivot_longer(
        cols = starts_with("week_"),
        names_to = "week", 
        values_to = "observation"
      )  |> 
      select(subject_id, arm, everything())
  })
```


Make a spaghetti plot showing observations on each subject over time.

```{r spaghetti_plot}
long_study_df |>
  ggplot(aes(x = week, y = observation, color = arm, group = subject_id)) +
  geom_line() +
  geom_point() +
  labs(title = "Observations Over Time") +
  facet_wrap(~arm) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


The plot shows us that the expiremental group has experienced a more significant increase in the observation values compared to the control group. The control group values were mostly consistent during the 8 week period.

### Problem 3

Let's set up the design elements for our one-sample t-test:

```{r power_simulation}
sim_power = function(mu) {
  
  data = tibble(
    x = rnorm(n = 30, mean = mu, sd = 5)
  )
  
  output = data |> 
    t.test() |> 
    broom::tidy() |> 
    select(estimate, p.value) |> 
    rename(mu_hat = estimate, p_val = p.value)
}

sim_power_results = expand_grid(
  mu_df = c(0, 1, 2, 3, 4, 5, 6), 
  iteration = 1:500
) |> 
  mutate(
    estimate = map(mu_df, sim_power)
  ) |> 
  unnest(estimate)
```


Now, let's plot the results of the simulation. 

First, we'll plot the proportion of times the null was rejected (in other words, the power of the test)

```{r power_plot}
sim_power_results |> 
  group_by(mu_df) |> 
  summarize(
    reject = sum(p_val < 0.05), 
    proportion = reject / 500
  ) |> 
  ggplot(aes(x = mu_df, y = proportion)) +
  geom_line(alpha = 0.6, linewidth = 0.9) +
  scale_x_continuous(breaks = seq(0,6, 1)) +
  labs(title = "Power Simulation (1:5000)")
```


















